---
title: "SpermCountAnalysis"
author: "Julin Maloof"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lme4)
```

Read in the data.  While the data appears continuous it is really normalized counts and I am rounding it off and treating it as such.  Also, set WT as the baseline for comparisons.
```{r}
d <- read_csv("Sperm viability raw data.csv") %>%
  mutate(across(starts_with("sperm"), round, digits=0),
         genotype = factor(genotype, levels = c("WT", "CrLFY1OX", "CrLFY2OX", "Cross"))) 
```

For this data format I need the counts in a separate matrix
```{r}
counts <- d %>% select(starts_with("sperm")) %>% as.matrix()
```

Null model (no genotype effect)
```{r}
m1 <- glmer(counts ~ (1|line), family = "binomial", data = d)
summary(m1)
```

```{r}
predict(m1, type = "response")
```

test genotype effect
```{r}
m2 <- glmer(counts ~ genotype + (1|line), family = "binomial", data = d)
summary(m2)
```
No genotypes significantly different from WT

Compare the null and genotype models to get an overall p-value
```{r}
anova(m1, m2)
```


